\documentclass[letterpaper,10pt]{article}

\usepackage{graphicx}                                        
\usepackage{amssymb}                                         
\usepackage{amsmath}                                         
\usepackage{amsthm}                                          

\usepackage{alltt}                                           
\usepackage{float}
\usepackage{color}
\usepackage{url}

\usepackage{balance}
\usepackage[TABBOTCAP, tight]{subfigure}
\usepackage{enumitem}
\usepackage{pstricks, pst-node}

\usepackage{listings}
\usepackage{geometry}
\geometry{textheight=8.5in, textwidth=6in}
\usepackage{booktabs}

%random comment

\newcommand{\cred}[1]{{\color{red}#1}}
\newcommand{\cblue}[1]{{\color{blue}#1}}

\newcommand{\toc}{\tableofcontents}

%\usepackage{hyperref}

%pull in the necessary preamble matter for pygments output
\input{pygments.tex}

\parindent = 0.0 in
\parskip = 0.1 in

\begin{document}

%\tableofcontents

\begin{center}
\textsc{\LARGE CS 444 Group 4}\\

% Title
{ \huge \bfseries Project 1 \\}

% Author
\emph{Member:}\\
Arvind \textsc{Vepa}
Chauncey \textsc{YAN}
\end{center}


\section{Command log for running the initial kernel}
\subsection{Setup the environment variables}
\subsubsection{bourne based shells}
source /scratch/opt/environment-setup-i586-poky-linux
\subsubsection{tcsh/csh shell}
source /scratch/opt/environment-setup-i586-poky-linux.csh
\subsection{copy the starting kernel, driver file, and other related files}
cp -R /scratch/spring2015/files/ .
\subsection{run the initial version of qemu}
qemu-system-i386 -gdb tcp::5504 -S -nographic -kernel bzImage-qemux86.bin -drive file=core-image-lsb-sdk-qemux86.ext3,if=virtio -enable-kvm -net none -usb -localtime --no-reboot --append ``root=/dev/vda rw console=ttyS0 debug''
\subsection{Open gdb in a seperate terminal and boot qemu}
gdb\\
target remote : 5504\\
continue\\
\subsection{Log in with root and no password from the qemu terminal}
\section{Command log for running the yocto kernel within the repository}
\subsection{Create new repository}
git init
\subsection{Acquire Yocto 3.14 kernel and switch to v3.14.24 tag}
git clone git://git.yoctoproject.org/linux-yocto-3.14\\
cd linux-yocto-3.14\\
git checkout tags/v3.14.24\\
\subsection{Copy configuration file to linux-yocto}
cd ..\\
cp config-3.14.26-yocto-qemu linux-yocto-3.14/.config\\
\subsection{Build the new instance of kernel}
cd linux-yocto-3.14\\
make -j4 all\\
\subsection{Run the qemu with the new generated inmage}
qemu-system-i386 -gdb tcp::5504 -S -nographic -kernel arch/x86/boot/bzImage -drive file=/scratch/spring2015/cs444-group04/core-image-lsb-sdk-qemux86.ext3,if=virtio -enable-kvm -net none -usb -localtime --no-reboot --append "root=/dev/vda rw console=ttyS0 debug"
\subsection{Using gdb to boot qemu}
gdb\\
target remote : 5504\\
continue\\
\subsection{Log in with root and no password from the qemu terminal}

\section{Work Log}
\begin{tabular}{@{}ccc@{}}
\toprule
Date                & Person   & Task                                                      \\ \midrule
Wed Apr 1           & Arvind   & pthreads                                                  \\
Wed Apr 1           & Chauncy  & random number generator                                   \\
Thu Apr 2           & Arvind   & pthreads                                                  \\
Sun Apr 5           & Arvind   & Finished pthreads                                         \\
Tue Apr 7           & Arvind   & Multiple threads                                          \\
Tue Apr 7           & Chauncy  & random number generator                                   \\
Wed Apr 8           & Chauncy  & Finished random number generator                          \\
Fri Apr 10          & Arvind   & Finished integrating                                      \\
Sun Apr 12          & Arvind   & assist with group write-up                                \\
Sun Apr 12          & Chauncy  & set up tex file                                           \\
Mon Apr 13          & Chauncy  & style guide                                               \\
Mon Apr 13          & Arvind   & style guide                                               \\ \bottomrule
\end{tabular}
\end{table}



\section{Version control log}

\begin{tabular}{@{}ccc@{}}
\toprule
Date                & Author   & Message                                                   \\ \midrule
Fri Apr 10 21:22:11 & Arvind   & version 1                                                 \\
Fri Apr 10 23:12:42 & Arvind   & version 2                                                 \\
Fri Apr 10 23:23:42 & Arvind   & mersenne twister files                                    \\
Sun Apr 12 20:43:22 & Arvind   & rdrand test                                               \\
Sun Apr 12 20:52:57 & Arvind   & version 4 tests complete                                  \\
Sun Apr 12 21:55:43 & Chauncey & late adding latex example                                 \\
Sun Apr 12 21:57:24 & Chauncey & Working on the groupwriteup                               \\
Sun Apr 12 21:58:22 & Chauncey & Finished detecting rdrand support and using rdrand and mt \\ \bottomrule
\end{tabular}
\end{table}

\section{Concurrency}
In this solution, we create two named semaphores and then a number of pthreads, corresponding to the threads specified on the command line (the first number is the number of producer threads an the second number is the number of consumer threads). After these threads are created, the threads are joined, so when the threads are terminated, they are joined.\par

In the file, we have a static member that corresponds to the pthread\_mutex lock and we have the initial declaration of the consumer and producer method, along with the declaration for the struct that contains both the number that will be printed out as well as the amount of time that it will wait.\par

In order to check if the system support rdrand instruction of not, check\_rd() is called. It runs asm code "cpuid \%0”, which return the cpu architecture information to a variable in terms of binary. Specifically, CPUID was called with register EAX = 1. ECX will return a 32 bits data that include all cpu info. If bit 30 is return 1 then the system support rdrand.\par

When the system has rdrand support, rnd\_int() is going to generate an unsigned int using the rdrand method. The way to using rdrand is through asm code. it generate an unsigned long integer from cpu and store it in the variable.\par

When the system has no rdrand support, such as OS-class, rnd\_int() is going to generate an unsigned int using Mersenne Twister method. First, we need to initialize mt with a seed array. Then we  call genrand\_int32() to get a random unsigned integer.\par

Within the consumer\_problem method, there is a do-while loop that runs infinitely. As specified in the directions, if the buffer is empty, then sem\_wait is called on the semaphore and the thread waits. Once the producer\_problem method calls sem\_post from the other method after adding an item to the buffer will the semaphore be released.\par

After the semaphore is released, the pthread is locked, and then the thread sleeps for the time specified within the first element. The number that is assigned to the first element is then printed. After this item is “consumed,” the element is overwritten by all the previous elements within the buffer, and then the semaphore that is used by the producer problem in order to signify that the buffer is ready to add an additional item when filled is posted, so that the producer may add another item. After the first item is removed from the buffer, the pthread is unlocked, and the do-while loop runs for a second iteration, and all subsequent iterations.\par

For the producer\_problem method, there is a do-while loop that runs infinitely. As specified in the directions, if the buffer is full, then sem\_wait is called on the semaphore and the thread waits. Once the consumer\_problem method calls sem\_post from the other method after removing an item from the buffer will the semaphore be released.\par

After the semaphore is released, the random number generator is initialized and the producer thread sleeps for a time between 2-7 seconds. Then the pthread is locked. Afterwards, then the next empty element of the buffer is filled with the struct containing the two random numbers and then the semaphore used by the consumer problem in order to signify that the buffer is ready to remove an additional item when the buffer is empty is posted, so that the consumer may remove another item. After the item is added to the buffer, the pthread is unlocked, and the do-while loop runs for a second iteration, and all subsequent iterations.\par

\section*{Appendix 1: Source Code}

\lstinputlisting[language=C]{consumer_producer_sol.c}
%\input{consumer_producer_sol.c}
%\input{__mt19937ar.c.tex}

\end{document}
